---
title: "Phyloseq PreProcessing"
author: "Alicia McElwee"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.path = "../figures/02_ProProcessing") # Sends any figure output to this folder
```


# Load Libraries
```{r install-libraries}

library(devtools)

#install_github("joey711/phyloseq")
library(phyloseq)

library(tidyverse)

```

# Goals

Here, we will process the data into a phyloseq object.

Three input files:

- ASV table
- Taxonomy table
- Track Reads (metadata)

Then, we will remove the following:

1.) Mitochondria
2.) Chloroplasts
3.) Remove samples without "enough" reads

Finally, write a data file of phyloseq output.



# Load Data

## ASV Table
```{r load-ASV-table}
# First, will load ASV table
load("data/01_DADA2/ASV_counts.RData")

# Inspect asv_tab
head(asv_tab)[,1:5]

# Fix sample names
sample_names <- colnames(asv_tab) # makes vector of sample filenames 
samples_fixed <- sapply(strsplit(basename(sample_names), "_"), `[`,1)
head(samples_fixed)

# re-write ASV count file to fix names
colnames(asv_tab) <- samples_fixed
str(asv_tab)

```

## Taxonomy Table
``` {r load-tax-table}
tax_df <- read.table("data/01_DADA2/ASV_taxonomy.tsv", sep = "\t", skip = 1)
head(tax_df)

# Fix column names
colnames(tax_df) <- c("asv_names","Kingdom","Phylum","Class", "Order", "Family", "Genus", "Species", "ASV", "ASV_seq")
head(tax_df)

# Make tax table a matrxi, Phyloseq needs tax table to be a matrix, requires row and column names
tax_mat <- tax_df %>%
  tibble::column_to_rownames(., var = "asv_names") %>% # takes asv_names column and makes it row names 
  as.matrix()


```

## Track Reads Data
``` {r load-track-reads}
load("data/01_DADA2/track_read_counts.RData")

# Take a look at the data
head(track_counts_df)
dim(track_counts_df)

#  Load in metadata
metadata_df <- read.csv("data/metadata.csv")
view(metadata_df)
dim(metadata_df)
colnames(metadata_df)

# Merge metadata_df with track_reads_df
head(metadata_df)
head(track_counts_df) # Both share same names, so will use that to join
metadata_track_reads_df <- 
  metadata_df %>%
  left_join(., track_counts_df, by = "names") %>%
  dplyr::select(-X)

# Intuition Check
head(metadata_track_reads_df)

# Update row.names to be sample names
## Before
row.names(metadata_track_reads_df)

## Rewrite
row.names(metadata_track_reads_df) <- metadata_track_reads_df$names
  
## Check that it worked
row.names(metadata_track_reads_df)

# Intuition check
head(metadata_track_reads_df)
```


# Handoff to Phyloseq

Combine data into phyloseq object

```{r phyloseq-handoff}
# Double check its all good
dim(asv_tab)
dim(tax_mat)

# Intuition check
stopifnot(row.names(asv_tab) == row.names(tax_mat)) # Code stops if what is in parenthesis is not true
    # If no output, is good!

# Construct phyloseq object
raw_physeq <- phyloseq(otu_table(asv_tab, taxa_are_rows = TRUE), 
                       sample_data(metadata_track_reads_df),
                       tax_table(tax_mat))
raw_physeq

# Save this raw phyloseq object
save(raw_physeq, file = "data/02_PreProcessing/raw_physeq.RData")
```

# Clean up the data

Remove:

1. Chloroplast
2. Mitochondria

```{r rm-mitos-chloros}
# Remind myself of tax table
view(tax_mat)

# Make new physeq without chloroplasts
noChloros_physeq <-
  raw_physeq %>%
  # rm chloroplasts
  subset_taxa(Order != "Chloroplast" | is.na(Order))

# How many taxa were chloroplasts?
num_chloro_ASVs <- ntaxa(raw_physeq) - ntaxa(noChloros_physeq)
num_chloro_ASVs

# Intuition check, checking that tax table of filtered data does not contain chloroplasts
#noChloros_physeq %>%
#  tax_table() %>%
#  data.frame() %>%
#  view()

# remove mitochondria
noChlorosMitos_physeq <- noChloros_physeq %>%
  subset_taxa(Family != "Mitochondria" | is.na(Family))

# How many mitochondrial ASVs?
num_mito_ASVs <-ntaxa(noChloros_physeq) - ntaxa(noChlorosMitos_physeq)
num_mito_ASVs

noChlorosMitos_physeq

# How many total ASVs were removed from Chloros and Mitos
total_ASVs_rm <- ntaxa(raw_physeq) - ntaxa(noChlorosMitos_physeq)
total_ASVs_rm

# Proportion ASVs kept?
ntaxa(noChlorosMitos_physeq)/ntaxa(raw_physeq)

```

# Evaluate and remove the control samples

Take a look at the negative controls and then make a decision about whether or not to remove the ASVS we found in our controls

1. Negative Controls
2. ASVs found in negative control and their distribution in the sample
3. Evaluate the mock community

## Negative Controls
```{r neg-controls}

# create a vector of samples that were negative controls
control_samples <- c("WaterControl", "022um-Control",
                     "3um-Control", "DNA-Ext-Control")

# Make a new phyloseq object of only the control samples
control_physeq <-
  noChlorosMitos_physeq %>%
  subset_samples(., (names %in% control_samples)) %>% # Found that controls had as many taxa as other samples
  #force remove ASVs with 0 counts
  prune_taxa(taxa_sums(.) > 0, .) # Get rid of 0 count ASVs in controls

control_physeq

# Are these 31 ASVs contaminates?

# Vector of control ASVs
control_ASVs <- 
  control_physeq %>%
  tax_table() %>%
  data.frame() %>%
  dplyr::select(ASV) %>%
  as.vector()

control_ASVs

# Visualize raw abundance values of control ASVs

control_physeq %>%
  plot_bar(., "ASV", fill = "ASV")
    # ASV 3 is most abundant, so 3rd most abundant ASV is in negative controls

# What is distribution of negative control ASVs between controls and actual samples?

######## Check the control ASVs in the sample
# Make new phyloseq object with only the controls ASVs in ALL samples
controlASV_physeq <- noChlorosMitos_physeq %>%
  subset_taxa(., ASV %in% control_ASVs$ASV)

# Inspect it
controlASV_physeq

# Take a look at the abundances
controlASV_physeq %>%
  plot_bar(., "ASV", fill = "ASV", facet_grid = Sample_or_Control~.)

  # Check ASVs that are in controls? If very abundant in controls and found in sample, likely want to remove them. Could do double check of taxonomy table of the ASVs to see what they may be/ are they likely to be contamination. Could also subtract out mean counts from controls

# Check tax of ASV_3
#control_physeq %>%
#  tax_table() %>%
#  data.frame() %>%
#  view()

# Would only like to keep ASV_18 and ASV_56, everything else is trash
controlASVs_toREMOVE <- dplyr::setdiff(control_ASVs$ASV, c("ASV_18","ASV_56"))
# double check
length(controlASVs_toREMOVE)

noControls_ChlorosMitos_physeq <-
  noChlorosMitos_physeq %>%
  # Actually remove the control ASVs (except 18, 56)
  subset_taxa(., !(ASV %in% controlASVs_toREMOVE)) %>%
  # REmove control samples
  subset_samples(., !(names %in% control_samples))
  
noControls_ChlorosMitos_physeq

# After filtering, how many ASVs were removed?
total_ASVs_rm <- ntaxa(raw_physeq) - ntaxa(noControls_ChlorosMitos_physeq)
total_ASVs_rm

```


# Session Information
```{r session-info}
# Ensure reproducibility by telling software versions
devtools::session_info()

```
